---
title: "Transfer subset of ADAM dataset and columns to RDF using D2RQ"
author: "mja@statgroup.dk"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    theme: united
  pdf_document:
    toc: true
    highlight: zenburn
  md_document:
    variant: markdown_github
---

# Introduction
This script creates RDF version of a subset of ADAM datasets and columns for the RDF data cubes.

# Setup

```{r}
library("rrdf")
library("SASxport")
library("RMySQL")
```


```{r}
targetDir<- "../res-ttl"
ttlDir<- "../res-ttl"
cdiscpilot01mapttlFn<- "cdiscpilot01-d2rq-mapping.ttl"
cdiscpilot01mapttlPath<- file.path(tempdir(), cdiscpilot01mapttlFn)
cdiscpilot01ttlFn<- "cdiscpilot01.ttl"
cdiscpilot01ttlPath<- file.path(tempdir(), cdiscpilot01ttlFn)
rqDir<- "../sparql-rq"
ADAMxptdir<-"../../phuse-scripts/data/adam/cdisc"
```

# Show parameters
```{r, eval=TRUE}
param.df<- data.frame(filepathname=apply(array(c(targetDir, ttlDir, rqDir, cdiscpilot01mapttlPath, cdiscpilot01ttlPath, ADAMxptdir)),1,normalizePath,mustWork = FALSE) )
rownames(param.df)<-c("Target diretory", ".ttl files for data cubes", "SPARQL scripts dir", "D2RQ mapping file for cdiscpilot01 data", "D2RQ file for cdiscpilot01 data", "Directory with ADAM xpt files" )
knitr::kable(param.df)
```

The following use of passwords in plain text is not recommeded.

MariaCB and R documentation Doc suggest to connect to my-db as defined in ~/.my.cnf.
See (https://github.com/rstats-db/RMySQL), (https://mariadb.com/kb/en/mariadb/configuring-mariadb-with-mycnf/), (http://stackoverflow.com/questions/2482234/how-to-know-mysql-my-cnf-location).

Before executing the commands below, MariaDB must be started. For MariaDB on fedora see (https://fedoraproject.org/wiki/MariaDB):

    systemctl start mariadb

## Identifying datasets and columns used in RDF data cubes

```{r, eval=TRUE}
store <- new.rdf(ontology=FALSE)  # Initialize
ttlFiles<- list.files( ttlDir, pattern="(TAB|FIG).+.ttl$")
for (dataCubeFile in ttlFiles) {
    dataCubeFn<- file.path(ttlDir, dataCubeFile )
    cat("Loading cube from ", dataCubeFn, "\n")
    tmp<- load.rdf(dataCubeFn, format="TURTLE", appendTo= store)
}

```

```{r, eval=TRUE}
rqFile<- file.path( rqDir, "variables-used-in-data-cubes.rq")
rq<- paste0(readLines(rqFile),collapse="\n")
res<- data.frame(sparql.rdf(store, rq),stringsAsFactors=FALSE)
print(nrow(res))
if (nrow(res)>0) {
    print(knitr::kable(res))
}
```

```{r, eval=TRUE}
parts1 <- data.frame(t(apply( res, 1, function(x) {
    m <- regexec("vocab/([^_]+)_([^_]+)$", x)
    (regmatches( x, m))[[1]][c(2,3)]
})), stringsAsFactors=FALSE)
colnames(parts1)<- c("dataset", "column")


```

Next step is to add the variables for the primary key. This is
currently done by hardcoding.  It can be automated by extracting the
key fields from `define-xml`. However, manual tweaking or complex
programming is needed, as some of the specified key variables have
missing values, which is not possible for variables being part of a
primary key in mariadb
(https://mariadb.com/kb/en/mariadb/primary-keys-with-nullable-columns/).

```{r, eval=TRUE}
adamkeys<- data.frame(
    matrix(c(
               "ADSL", "USUBJID",
               "ADAE", "USUBJID",
               "ADAE", "AESEQ",
               "ADQSADAS", "USUBJID",
               "ADQSADAS", "PARAMCD",
               "ADQSADAS", "AVISIT",
               "ADQSADAS", "ADT"
           ), ncol=2, nrow=7, byrow=TRUE )
  , stringsAsFactors=FALSE)
colnames(adamkeys)<- c("dataset", "column")
adamkeys
parts2<- rbind( parts1, adamkeys )
               
```

Finally, show the variables.
```{r, eval=TRUE}
parts3<- parts2[order(parts2[,1],parts2[,2]),]
parts<- parts3[!duplicated(parts3),]
row.names(parts)<- paste(seq(nrow(parts)))
knitr::kable(parts)
```

## Converting .xpt files to mariadb

The R to mariadb conversion needs to specify data types, specifically, the length of strings and date and data time objects.
This is done variable by variable. The approach is easier to debug than doing it per dataset (using `apply`).

```{r, eval=TRUE}
con <- dbConnect(RMySQL::MySQL(), username="root", password="mariadb")
res <- dbSendQuery(con, paste( "DROP  DATABASE IF EXISTS cdiscpilot01;"))
res <- dbSendQuery(con, paste( "CREATE DATABASE cdiscpilot01;"))
dbDisconnect(con)
con <- dbConnect(RMySQL::MySQL(), username="root", password="mariadb", dbname="cdiscpilot01" )
dbListTables(con)
dsname<- "."
i<- 1
for (i in seq(nrow(parts))) {
    if (parts$dataset[i] != dsname ) {
        dsname<- parts$dataset[i]
        xpt.fn<- file.path( ADAMxptdir, paste0(tolower(dsname), ".xpt"))
        cat("Reading ", xpt.fn, "\n")
        adds.full<- NULL
        adds.xfer<- NULL
        adds.full<- SASxport::read.xport(xpt.fn, as.is=TRUE)
        adds.def<- lookup.xport(xpt.fn)[[dsname]]
##        adds.xfer<- adds.full
        adds.xfer.columns<- parts[parts$dataset==dsname,"column"]
        field.type.def<- list(  )
        ii<- 1
  ##      adds.xfer<- adds.full[,adds.def$name %in% adds.xfer.columns]
      adds.xfer<- adds.full[,adds.def$name %in% adds.xfer.columns]
    ## str(adds.full)
    str(adds.xfer)
    }
    vn<- parts$column[i]
    if (!any(adds.def$name==vn)) {
        cat("--> ", vn, " not found in ", dsname, "\n")
    } else {
        vtype<- adds.def$type[adds.def$name==vn]
        vformat<- adds.def$format[adds.def$name==vn]
        cat(vn, vtype, vformat, "\n")
        if (is.character(vtype) && is.character(vformat) ) {
            if ( vtype=="numeric" ) {
                if (vformat=="DATE") {
                    adds.xfer[,vn]<- as.Date(adds.xfer[,vn])
                    field.type.def[[vn]]<- "date"
                } else  { ## add more, consider using switch
                    field.type.def[[vn]]<- "double"                    
                }
            } else if ( vtype=="character" ) {
                vlen<- max(nchar(adds.xfer[,vn]))
                field.type.def[[vn]]<- paste0("varchar","(", vlen, ")")
            } else {
                cat("--> ", "Unhandled: ", vn, vtype, vformat, "\n")
            }
        }
    }
    if (i == nrow(parts) | i < nrow(parts) & parts$dataset[i+1] != dsname ) {
        dbWriteTable(con, dsname, adds.xfer, field.type=field.type.def, overwrite=TRUE, row.names=FALSE)
        sqlcmd<- paste0("ALTER TABLE ", dsname,
                        " ADD PRIMARY KEY", "(",
                        paste0(adamkeys[adamkeys$dataset==dsname,"column"],collapse=", "),
                        ")",
                        collapse=" " )
        cat("sql cmd: ", sqlcmd, "\n")        
       res <- dbSendQuery(con, sqlcmd )
        cat("Done\n")
    }
}          
dbDisconnect(con)

```

# Use D2RQ to transfer to .ttl files

d2rq is used to convert the database.
First a mapping file is generated and customized.
The `d2rq:classDefinitionLabel` could also be changed to correspond to the SAS label.

```{r, eval=TRUE}
d2rqbaseURL<- "http://www.example.org/datasets/"
d2rqjdbcstring<- '"jdbc:mariadb://localhost:3306/cdiscpilot01?user=root&password=mariadb"'
## -b option does not work with -l
## /opt/d2rq-0.8.1/generate-mapping reports Unknown argument: -b
system(
    paste("/opt/d2rq-0.8.1/generate-mapping",
          " -o ", cdiscpilot01mapttlPath,
          " --verbose",
             " ", d2rqjdbcstring, " "
          )
)

d2rqmap<- readLines(cdiscpilot01mapttlPath)
if (d2rqmap[1]== "@prefix map: <#> .") {
    d2rqmap[1]<- "@prefix map: <http://www.example.org/db/map/> ."
}
if (d2rqmap[2]== "@prefix db: <> .") {
    d2rqmap[2]<- "@prefix db: <http://www.example.org/db/> ."
}    
if (d2rqmap[3]== "@prefix vocab: <vocab/> .") {
    d2rqmap[3]<- "@prefix vocab: <http://www.example.org/datasets/vocab/> ."
}
fp<- file.path(targetDir, cdiscpilot01mapttlFn)
writeLines(d2rqmap, fp )
cat( "File ", fp, " created", "\n")

```


```{r, eval=TRUE}
system(paste("/opt/d2rq-0.8.1/dump-rdf",
             " -b ", d2rqbaseURL,
             " --verbose",
             " -o ", cdiscpilot01ttlPath,
             " ", cdiscpilot01mapttlPath, " "
             ))
if (file.copy(cdiscpilot01ttlPath, targetDir, overwrite = TRUE)) {
    cat( "File ", cdiscpilot01ttlPath, " copied to directory ", targetDir, "\n")
}

```


# Configuration of mariadb


## If RMySQL gives an errormessage

After installation of mariadb

    install.packages("RMySQL")

gives error

    libmariadb.so.2: cannot open shared object file: No such file or directory

but the files exists as 

    /usr/lib64/mariadb/libmariadb.so.2

and the prerequisites in https://cran.r-project.org/web/packages/RMySQL/README.html are met.

My solution was to modify '/etc/ld.so.conf.d/mariadb-x86_64.conf' by adding a reference to

    /usr/lib64/mariadb


## Installing MariaDB on fedora 

Follow the instructions at (https://fedoraproject.org/wiki/MariaDB).

For my setup I start mariadb for each session by

    systemctl start mariadb

Alternatively, could be to have mariadb running as service.

## D2RQ jdbc and mariadb 

For d2rq jdbc - get mariadb-java-client from (). I used
`mariadb-java-client-1.5.2.jar` and store it in
`/opt/d2rq-0.8.1/lib/db-drivers/`.

# Final stuff

Here is how to make the PDF file
```{r, eval=FALSE}
rmarkdown::render('ADAM-ds-as-ttl-using-d2rq.Rmd', c("html_document"), clean=TRUE)
```

