--
title: "Get data for RDF data cube observation"
author: "mja@statgroup.dk"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    theme: united
  pdf_document:
    toc: true
    highlight: zenburn
  md_document:
    variant: markdown_github
---

# Introduction
Here I show how to get the data for for a statistics in a RDF data cube observation.

# Setup

## Loading libraries

```{r}
library(rrdfancillary)
library(rrdfcdisc)
library(rrdfqb)
library(rrdfqbcrnd0)
library(knitr)

```

## Internals

The display of SPARQL script in markdown is done by first creating a
chunk, and then using the chunk with the highlight engine in
knitr. The advantage of this approach is that all formatting is
handled by external packages. To make the highlight output work in
markdown two blanks has to be added at the end of line according to
markdown syntax. The function stores the sparql statements in a temporary file, which then is processed by `knitr::read_chunk`.
The pdf generation with `rmarkdown::render` did not handle backslash-n (newline), so to avoid writing it I define the  as a constant `const.newline`

```{r helperfunction}
mdwrite<- function( sparqlStatements, refname ) {
fn<- file.path(tempdir(), paste0( refname, ".rq" ) )
cat( paste0("## @knitr ", refname), gsub("\\n", "  \n", sparqlStatements), sep="  \n", file=fn)
knitr::read_chunk( fn, from=c(1))
invisible(fn)
}

const.newline<- "\n"
```

Here is an example of the R code to make the chunk:
```{r, eval=FALSE}
mdwrite( dimensionsRq, "DEMOdimensions" )

```

The chunk is then shown using

```
    {r DEMOdimensions, results='asis', engine='highlight', engine.opts='-S n3 --inline-css'}
       
```

# Setup

First, I define the location for the RDF data cube
```{r}
rdf.data.cube.Name<- "TAB2X01"
rdf.data.cube.Dir<- "../res-ttl"
rdf.data.cube.File<- file.path(rdf.data.cube.Dir, paste("CDISC-pilot-", rdf.data.cube.Name, ".ttl", sep=""))

```

Second, I define the location for the `ADSL` data as turtle transformed using R2DQ (see `rrdfqbcrnd0/rrdfqbcrndex/inst/data-raw/create-ADSL-ttl.Rmd`).

```{r}
dataFilemap<- system.file("extdata/sample-rdf", "adsl-map.ttl", package="rrdfqbcrndex")


```

The final step is to load the turtle files into a RDF store, here named `storeCube`.

```{r}
storeCube <- new.rdf(ontology=FALSE)  
temp<- load.rdf(rdf.data.cube.File, format="TURTLE", appendTo= storeCube)
cat("Reading RDF Data Cube from file ", normalizePath(rdf.data.cube.File), const.newline)
summarize.rdf(storeCube)
cat("Reading data set D2RQ map from file ", normalizePath(dataFilemap), const.newline)
temp<- load.rdf(dataFilemap, format="TURTLE", appendTo= storeCube)
dataFile<- system.file("extdata/sample-rdf", "adsl.ttl", package="rrdfqbcrndex")
cat("Reading data set in D2RQ format from file ", normalizePath(dataFile), const.newline)
temp<- load.rdf(dataFile, format="TURTLE", appendTo= storeCube)
summarize.rdf(storeCube)

```

Now, just to see if it works, I will get the data for one record. The key URI for a record is composed by the name of the dataset `ADSL` and the `usubjid` value `01-718-1254`.

```{r}
d2rqbaseURL<- "http://www.example.org/datasets/"
d2rqvocab<- paste0(d2rqbaseURL, "vocab", "/", sep="")

s<- paste0("<", d2rqbaseURL,c("ADSL/01-718-1254"), ">")
data.records.rq<-paste(
    "select * ",
    "where { ?s ?p ?o.",
    " values(?s) {",
    paste("(",s,")", collapse=const.newline),
    "}",
    "}", sep=const.newline, collapse=const.newline )
mdwrite( data.records.rq, "ADSL-records-SELECT-query.rq" )
records.res<- data.frame(sparql.rdf(storeCube, data.records.rq),stringsAsFactors=FALSE)

```

The SPARQL query is

```{r ADSL-records-SELECT-query.rq, results='asis', engine='highlight', engine.opts='-S n3 --inline-css'}
```    

The result is

```{r results='asis', engine='highlight', engine.opts='-S n3 --inline-css'}
cat(paste(apply(records.res,1,FUN=paste0, collapse=" "),collapse=const.newline),const.newline)
```    


```{r}

## Get the values in the cube
dsdName<- GetDsdNameFromCube( storeCube )
domainName<- GetDomainNameFromCube( storeCube )
forsparqlprefix<- GetForSparqlPrefix( domainName )
## Get cube components
componentsRq<- GetComponentSparqlQuery( forsparqlprefix, dsdName )
components<- as.data.frame(sparql.rdf(storeCube, componentsRq), stringsAsFactors=FALSE)
components$vn<- gsub("crnd-dimension:|crnd-attribute:|crnd-measure:","",components$p)
knitr::kable(components[,c("vn", "label")])
## Get code lists
codelistsRq<- GetCodeListSparqlQuery( forsparqlprefix, dsdName )
codelists<- as.data.frame(sparql.rdf(storeCube, codelistsRq), stringsAsFactors=FALSE)
codelists$vn<- gsub("crnd-dimension:|crnd-attribute:|crnd-measure:","",codelists$dimension)
codelists$clc<- gsub("code:","",codelists$cl)
knitr::kable(codelists[,c("vn", "clc", "clprefLabel")])
## Get dimensions
dimensionsRq <- GetDimensionsSparqlQuery( forsparqlprefix )
dimensions<- sparql.rdf(storeCube, dimensionsRq)
knitr::kable(dimensions)
## Get attributes
attributesRq<- GetAttributesSparqlQuery( forsparqlprefix )
attributes<- sparql.rdf(storeCube, attributesRq)
knitr::kable(attributes)
```
The result provides for each observation the data values that must match.

## Create the SPARQL query

### Step 1: for an observation find the values that must match

```{r}
qobs<- "ds:obs223"


dobs.dimensions.values.rq<- paste(
    "select * where {",
"
   ?obs ?dim ?codevalue .
   ?dim a qb:DimensionProperty .
   ?codelist skos:hasTopConcept ?codevalue .
   ?codelist rrdfqbcrnd0:DataSetRefD2RQ ?vnop . 
   ?codelist rrdfqbcrnd0:R-columnname ?vn . 
   ?codelist rrdfqbcrnd0:codeType     ?vct .         
   ?codevalue skos:prefLabel ?clprefLabel . 
   ?codevalue rrdfqbcrnd0:R-selectionoperator ?Rselectionoperator . 
   ?codevalue rrdfqbcrnd0:R-selectionvalue ?Rselectionvalue .   
",
"values (?obs) {", paste0("(", qobs, ")",collapse=const.newline), "}",
   "}"
)                                  
dobs.rq<-     paste( forsparqlprefix, dobs.dimensions.values.rq )
# xx
dobsobs<- NULL
dobsobs<- as.data.frame(sparql.rdf(storeCube, dobs.rq ), stringsAsFactors=FALSE)
knitr::kable(dobsobs)

```

For subsequent use, the query is changed to only select the variables need for getting the data.

```{r}
dobs1.dimensions.values.rq<- paste(
    "select ?vnop ?Rselectionvalue",
    " (replace(str(?vnop),'http://www.example.org/rrdfqbcrnd0/([A-Z0-9_]+)$', '$1', 'i' ) as ?d2rqname)\n",
    paste0(" ( concat(", "'", d2rqvocab, "'", ", ", "replace(str(?vnop),'http://www.example.org/rrdfqbcrnd0/([A-Z0-9_]+)$', '$1', 'i' )) as ?d2rqIRI)\n"),
    "where {",
"
   ?obs ?dim ?codevalue .
   ?dim a qb:DimensionProperty .
   ?codelist skos:hasTopConcept ?codevalue .
   ?codelist rrdfqbcrnd0:DataSetRefD2RQ ?vnop . 
   ?codelist rrdfqbcrnd0:R-columnname ?vn . 
   ?codelist rrdfqbcrnd0:codeType     ?vct .         
   ?codevalue skos:prefLabel ?clprefLabel . 
   ?codevalue rrdfqbcrnd0:R-selectionoperator ?Rselectionoperator . 
   ?codevalue rrdfqbcrnd0:R-selectionvalue ?Rselectionvalue .   
",
"values (?obs) {", paste0("(", qobs, ")",collapse=const.newline), "}",
   "}"
)                                  
dobs1.rq<-     paste( forsparqlprefix, dobs1.dimensions.values.rq )
## xx
dobs1obs<- NULL
dobs1obs<- as.data.frame(sparql.rdf(storeCube, dobs1.rq ), stringsAsFactors=FALSE)
knitr::kable(dobs1obs)

```

For subsequent use, the query is changed to only select the variables need for getting the data.

 

```{r}
dobs1.dimensions.values.rq<- paste(
    "select ",
    paste0(" ( iri(concat(", "'", d2rqvocab, "'", ", ", "replace(str(?vnop),'http://www.example.org/rrdfqbcrnd0/([A-Z0-9_]+)$', '$1', 'i' ))) as ?variable)\n"),
    "?matchvalue\n",
    "where {",
"
   ?obs ?dim ?codevalue .
   ?dim a qb:DimensionProperty .
   ?codelist skos:hasTopConcept ?codevalue .
   ?codelist rrdfqbcrnd0:DataSetRefD2RQ ?vnop . 
   ?codelist rrdfqbcrnd0:R-columnname ?vn . 
   ?codelist rrdfqbcrnd0:codeType     ?vct .         
   ?codevalue skos:prefLabel ?clprefLabel . 
   ?codevalue rrdfqbcrnd0:R-selectionoperator ?Rselectionoperator . 
   ?codevalue rrdfqbcrnd0:R-selectionvalue ?matchvalue.
",
"values (?obs) {", paste0("(", qobs, ")",collapse=const.newline), "}",
   "}"
)                                  
dobs1.rq<-     paste( forsparqlprefix, dobs1.dimensions.values.rq )
## xx
dobs1obs<- NULL
dobs1obs<- as.data.frame(sparql.rdf(storeCube, dobs1.rq ), stringsAsFactors=FALSE)
knitr::kable(dobs1obs)

```


### Step 2 


Next step is to retrieve the data, using the values hardcoded.
The next query shows the matching records in the usual rows and columns format, with the columns represeting the unique key for the record, defined by D2RQ, and the two columns in question `TRT01P` and `ITTFL`.

```{r}

records.rq<- paste("select * ",
           "where {
?s d2rqvocab:ADSL_TRT01P ?TRT01P .
?s d2rqvocab:ADSL_ITTFL ?ITTFL .
?s d2rqvocab:ADSL_TRT01P 'Xanomeline High Dose' ;
   d2rqvocab:ADSL_ITTFL  'Y' .
",
"}",
"order by ?s",
sep=const.newline, collapse=const.newline )
dobsds.rq<-     paste( forsparqlprefix,
                      paste0( "prefix d2rqvocab: ", "<", d2rqvocab, ">", collapse=""),
                      records.rq )
## xx
dobsobs<- NULL
dobsobs<- as.data.frame(sparql.rdf(storeCube, dobsds.rq ), stringsAsFactors=FALSE)
knitr::kable(dobsobs)

```

Another representation is showing the the variables one in each row.

```{r}

obs.triples1.rq<-paste("select * ",
           "where {
?s ?variable ?value .
values (?variable) {
(d2rqvocab:ADSL_TRT01P)
(d2rqvocab:ADSL_ITTFL)
}
",
"}",
"order by ?s",
sep=const.newline, collapse=const.newline )
dobs.triples1.ds.rq<-     paste( forsparqlprefix,
                      paste0( "prefix d2rqvocab: ", "<", d2rqvocab, ">", collapse=""),
                      obs.triples1.rq )
## xx
dobs.triples1.obs<- NULL
dobs.triples1.obs<- as.data.frame(sparql.rdf(storeCube, dobs.triples1.ds.rq ), stringsAsFactors=FALSE)
knitr::kable(head(dobs.triples1.obs,10))

```

Next step is adding the matching value and reporting if its equal with 0 and 1.
Note the BIND is after the values statement.

```{r}

obs.triples2.rq<-paste("select * ",
           "where {
?s ?variable ?value .
values (?variable ?matchvalue) {
(d2rqvocab:ADSL_TRT01P 'Xanomeline High Dose')
(d2rqvocab:ADSL_ITTFL 'Y' )
}
BIND(IF(?value=?matchvalue,1,0) AS ?isequal)
",
"}",
"order by ?s",
sep=const.newline, collapse=const.newline )
dobs.triples2.ds.rq<-     paste( forsparqlprefix,
                      paste0( "prefix d2rqvocab: ", "<", d2rqvocab, ">", collapse=""),
                      obs.triples2.rq )
## xx
dobs.triples2.obs<- NULL
dobs.triples2.obs<- as.data.frame(sparql.rdf(storeCube, dobs.triples2.ds.rq ), stringsAsFactors=FALSE)
knitr::kable(head(dobs.triples2.obs,10))

```

Using SPARQL aggregate query grouping by `?s` provides the number of not-matching values.
Now, the desired records `?s` are those where there is 0 not-equal variables.

```{r}

obs.triples3.rq<-paste("SELECT ?s (SUM(?notequal) as ?nnotqual) ",
           "where {
?s ?variable ?value .
values (?variable ?matchvalue) {
(d2rqvocab:ADSL_TRT01P 'Xanomeline High Dose')
(d2rqvocab:ADSL_ITTFL 'Y' )
}
BIND(IF(?value!=?matchvalue,1,0) AS ?notequal)
",
"}",
"group by ?s",
"order by ?s",
sep=const.newline, collapse=const.newline )
dobs.triples3.ds.rq<-     paste( forsparqlprefix,
                      paste0( "prefix d2rqvocab: ", "<", d2rqvocab, ">", collapse=""),
                      obs.triples3.rq )
## xx
dobs.triples3.obs<- NULL
dobs.triples3.obs<- as.data.frame(sparql.rdf(storeCube, dobs.triples3.ds.rq ), stringsAsFactors=FALSE)
knitr::kable(head(dobs.triples3.obs))

```

Then to get the desired records, identified by `?s`, the SPARQL HAVING term is used.

```{r}

obs.triples4.rq<-paste("SELECT ?s ",
           "where {
?s ?variable ?value .
values (?variable ?matchvalue) {
(d2rqvocab:ADSL_TRT01P 'Xanomeline High Dose')
(d2rqvocab:ADSL_ITTFL 'Y' )
}
BIND(IF(?value!=?matchvalue,1,0) AS ?notequal)
",
"}",
"group by ?s",
"having(SUM(?notequal)=0)",
"order by ?s",
sep=const.newline, collapse=const.newline )
dobs.triples4.ds.rq<-     paste( forsparqlprefix,
                      paste0( "prefix d2rqvocab: ", "<", d2rqvocab, ">", collapse=""),
                      obs.triples4.rq )
## xx
dobs.triples4.obs<- NULL
dobs.triples4.obs<- as.data.frame(sparql.rdf(storeCube, dobs.triples4.ds.rq ), stringsAsFactors=FALSE)
knitr::kable(head(dobs.triples4.obs))

```

The query above can then be used as a subquery in the full querqy.

```{r}

records2.rq<-paste("select * ",
           "where {
?s d2rqvocab:ADSL_TRT01P ?TRT01P .
?s d2rqvocab:ADSL_ITTFL ?ITTFL .
",
"{",
obs.triples4.rq,
"}",
"}",
"order by ?s",
sep=const.newline, collapse=const.newline )

dobs2ds.rq<-     paste( forsparqlprefix,
                      paste0( "prefix d2rqvocab: ", "<", d2rqvocab, ">", collapse=""),
                      records2.rq )
## xx
dobs2obs<- NULL
dobs2obs<- as.data.frame(sparql.rdf(storeCube, dobs2ds.rq ), stringsAsFactors=FALSE)
knitr::kable(dobs2obs)

```

The values part of the query above can be replaced by a subquery.

```{r}


obs.triples5.rq<-paste(
paste(
    "select ",
    paste0(" ( iri(concat(", "'", d2rqvocab, "'", ", ", "replace(str(?vnop),'http://www.example.org/rrdfqbcrnd0/([A-Z0-9_]+)$', '$1', 'i' ))) as ?variable)\n"),
    "?matchvalue\n",
    "where {",
"
   ?obs ?dim ?codevalue .
   ?dim a qb:DimensionProperty .
   ?codelist skos:hasTopConcept ?codevalue .
   ?codelist rrdfqbcrnd0:DataSetRefD2RQ ?vnop . 
   ?codelist rrdfqbcrnd0:R-columnname ?vn . 
   ?codelist rrdfqbcrnd0:codeType     ?vct .         
   ?codevalue skos:prefLabel ?clprefLabel . 
   ?codevalue rrdfqbcrnd0:R-selectionoperator ?Rselectionoperator . 
   ?codevalue rrdfqbcrnd0:R-selectionvalue ?matchvalue.
",
"values (?obs) {", paste0("(", qobs, ")",collapse=const.newline), "}",
"}"
),
sep=const.newline, collapse=const.newline )
dobs.triples5.ds.rq<-     paste( forsparqlprefix,
                      paste0( "prefix d2rqvocab: ", "<", d2rqvocab, ">", const.newline, collapse=""),
                      obs.triples5.rq )
## xx
dobs.triples5.obs<- NULL
dobs.triples5.obs<- as.data.frame(sparql.rdf(storeCube, dobs.triples5.ds.rq ), stringsAsFactors=FALSE)
knitr::kable(head(dobs.triples5.obs))

obs.triples6.rq<-paste(
    "SELECT ?s",
           "where {
?s ?variable ?value 
",
"{",
paste(
    "select ",
    paste0(" ( iri(concat(", "'", d2rqvocab, "'", ", ", "replace(str(?vnop),'http://www.example.org/rrdfqbcrnd0/([A-Z0-9_]+)$', '$1', 'i' ))) as ?variable)\n"),
    "?matchvalue\n",
    "where {",
"
   ?obs ?dim ?codevalue .
   ?dim a qb:DimensionProperty .
   ?codelist skos:hasTopConcept ?codevalue .
   ?codelist rrdfqbcrnd0:DataSetRefD2RQ ?vnop . 
   ?codelist rrdfqbcrnd0:R-columnname ?vn . 
   ?codelist rrdfqbcrnd0:codeType     ?vct .         
   ?codevalue skos:prefLabel ?clprefLabel . 
   ?codevalue rrdfqbcrnd0:R-selectionoperator ?Rselectionoperator . 
   ?codevalue rrdfqbcrnd0:R-selectionvalue ?matchvalue.
",
"values (?obs) {", paste0("(", qobs, ")",collapse=const.newline), "}",
"}"
),
"}",
"BIND(IF(?value!=?matchvalue,1,0) AS ?notequal)",
"}",
"group by ?s",
"having(SUM(?notequal)=0)",
"order by ?s",
sep=const.newline, collapse=const.newline )
dobs.triples6.ds.rq<-     paste( forsparqlprefix,
                      paste0( "prefix d2rqvocab: ", "<", d2rqvocab, ">", const.newline, collapse=""),
                      obs.triples6.rq )
## xx
dobs.triples6.obs<- NULL
dobs.triples6.obs<- as.data.frame(sparql.rdf(storeCube, dobs.triples6.ds.rq ), stringsAsFactors=FALSE)
knitr::kable(dobs.triples6.obs)


```

# Final stuff

Here is how to make the PDF file
```{r, eval=FALSE}
rmarkdown::render('get-data-for-cube-observation.Rmd', c("html_document", "pdf_document"), clean=TRUE)
```
