# Loading libraries

```{r}
library(rrdfancillary)
library(rrdfcdisc)
library(rrdfqb)
library(rrdfqbcrnd0)
library(knitr)
```

# Show parameters


```{r, eval=TRUE}
param.df<- data.frame(filepathname=apply(array(c(targetName, targetDir, targetFile, targetObsrqFile, target2DrqFile)),1,normalizePath,mustWork = FALSE) )
rownames(param.df)<-c("Target file name", "Target diretory", "Target File", "Observation query file", "Row-column query file")
knitr::kable(param.df)
```

# Load CSV files

```{r, eval=TRUE}
ObsDataCsvFn<- file.path("../res-csv", paste( targetName, ".csv", sep=""))
cat("Observations CSV file ", normalizePath(ObsDataCsvFn), "\n")
ObsData <- read.csv(ObsDataCsvFn,stringsAsFactors=FALSE)
MetaDataCsvFn<- file.path("../res-csv", paste( targetName, "-Components.csv", sep=""))
cat("MetaData CSV file ", normalizePath(MetaDataCsvFn), "\n")
MetaData <- read.csv(MetaDataCsvFn,stringsAsFactors=FALSE)
```

# Create cube

```{r, eval=TRUE}
cube.fn<- BuildCubeFromDataFrames(MetaData, ObsData )
cat(targetName, " ", "cube stored as ", normalizePath(cube.fn), "\n")
```
# Copy cube to destination directory

```{r, eval=TRUE}
if (file.copy( cube.fn, targetFile, overwrite=TRUE)) {
   cat("RDF data cube copied to ", normalizePath(targetFile), "\n")
 }
```

# Get SPARQL query for observations

This is lifted from the example rrdfqbcrndex/vignettes/cube-from-workbook.Rmd.

```{r, eval=TRUE}
checkCube <- new.rdf()  # Initialize
temp<- load.rdf(targetFile, format="TURTLE", appendTo= checkCube)
cat("Reading RDF Data Cube from file ", normalizePath(targetFile), "\n")
summarize.rdf(checkCube)
## Get the values in the cube
dsdName<- GetDsdNameFromCube( checkCube )
domainName<- GetDomainNameFromCube( checkCube )
forsparqlprefix<- GetForSparqlPrefix( domainName )
## Get cube components
componentsRq<- GetComponentSparqlQuery( forsparqlprefix, dsdName )
components<- as.data.frame(sparql.rdf(checkCube, componentsRq), stringsAsFactors=FALSE)
components$vn<- gsub("crnd-dimension:|crnd-attribute:|crnd-measure:","",components$p)
knitr::kable(components[,c("vn", "label")])
## Get code lists
codelistsRq<- GetCodeListSparqlQuery( forsparqlprefix, dsdName )
codelists<- as.data.frame(sparql.rdf(checkCube, codelistsRq), stringsAsFactors=FALSE)
codelists$vn<- gsub("crnd-dimension:|crnd-attribute:|crnd-measure:","",codelists$dimension)
codelists$clc<- gsub("code:","",codelists$cl)
knitr::kable(codelists[,c("vn", "clc", "clprefLabel")])
## Get dimensions
dimensionsRq <- GetDimensionsSparqlQuery( forsparqlprefix )
dimensions<- sparql.rdf(checkCube, dimensionsRq)
knitr::kable(dimensions)
## Get attributes
attributesRq<- GetAttributesSparqlQuery( forsparqlprefix )
attributes<- sparql.rdf(checkCube, attributesRq)
knitr::kable(attributes)
```

## Get observations
And finally the SPARQL query for observations, showing only the first 10 observations.
```{r, echo=TRUE}
observationsRq<- GetObservationsSparqlQuery( forsparqlprefix, domainName, dimensions, attributes )
cat(observationsRq)
cat(observationsRq, file=targetObsrqFile)
observations<- as.data.frame(sparql.rdf(checkCube, observationsRq ), stringsAsFactors=FALSE)
knitr::kable(observations[ 1:10 ,
   c(paste0(sub("crnd-dimension:|crnd-attribute:|crnd-measure:", "", dimensions), "value"),sub("crnd-dimension:|crnd-attribute:|crnd-measure:", "", attributes), "measure")])

```

## Get observations with labels

The SPARQL query for observations with labels for variables, showing only the first 10 observations.
```{r, echo=TRUE}
observationsDescriptionRq<- GetObservationsWithDescriptionSparqlQuery( forsparqlprefix, domainName, dimensions, attributes )
cat(observationsDescriptionRq)
observationsDesc<- as.data.frame(sparql.rdf(checkCube, observationsDescriptionRq ), stringsAsFactors=FALSE)
knitr::kable(observationsDesc[ 1:10 ,
     c(paste0(rep(sub("crnd-dimension:|crnd-attribute:|crnd-measure:", "", dimensions),each=3),
       c("label", "value", "IRI")),
       sub("crnd-dimension:|crnd-attribute:|crnd-measure:", "", attributes), "measure", "measureIRI"
       )]
       )

```

# Session information
```{r, echo=TRUE}
sessionInfo()
```
