---
title: "Create CDISC pilot tables"
author: "mja@statgroup.dk"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    theme: united
  pdf_document:
    toc: true
    highlight: zenburn
  md_document:
    variant: markdown_github
---


# Setup

# This script creates RDF data cubes from .csv files.


library(rrdfancillary)
library(rrdfcdisc)
library(rrdfqb)
library(rrdfqbcrnd0)
```

#All files are stored in the directory
#```{r}
targetDir<- "../res-ttl"
(targetDir)


# Create from demo data PhUSE scripting program

#The files are generated by a SAS program gen-res-sas/tab1x01.sas.  The input files are read
#directly from .csv files instead of using the workbook, to avoid the
#extra step with the workbook, as the input files may be updated by
#running the SAS program.

# {r, eval=TRUE}
tab1x03ObsDataCsvFn<- file.path("../res-csv", "TAB1X03.csv")
tab1x03ObsData <- read.csv(tab1x03ObsDataCsvFn,stringsAsFactors=FALSE)


tab1x03MetaDataCsvFn<- file.path("../res-csv", "TAB1X03-Components.csv")
tab1x03MetaData <- read.csv(tab1x03MetaDataCsvFn,stringsAsFactors=FALSE)


tab1x03.cube.fn<- BuildCubeFromDataFrames(tab1x03MetaData, tab1x03ObsData )
cat("TAB1X03 cube stored as ", normalizePath(tab1x03.cube.fn), "\n")

targetFile<- file.path(targetDir,"CDISC-pilot-TAB1X03.ttl")

if (file.copy( tab1x03.cube.fn, targetFile, overwrite=TRUE)) {
   cat("RDF data cube copied to ", normalizePath(targetFile), "\n")
 }


#Now look at the generated cubes by loading the turtle files. 

#{r, echo=TRUE, results='asis'}
dataCubeFile<- targetFile


#The rest of the code only depends on the value of dataCubeFile.
#```{r, echo=TRUE}
checkCube <- new.rdf(ontology=FALSE)  # Initialize
cat("Loading RDF data cube from ", normalizePath(dataCubeFile), "\n")
temp<- load.rdf(dataCubeFile, format="TURTLE", appendTo= checkCube)
summarize.rdf(checkCube)


## Get the values in the cube
##First set values for accessing the cube.
##{r, echo=TRUE}
dsdName<- GetDsdNameFromCube( checkCube )
domainName<- GetDomainNameFromCube( checkCube )
forsparqlprefix<- GetForSparqlPrefix( domainName )


#The next statement shows the first 10 triples in the cube.
#```{r, echo=TRUE}
observations1Rq<- paste( forsparqlprefix,
'
select *
where {?s ?p ?o .}
limit 100
',
"\n"
)
observations1<- sparql.rdf(checkCube, observations1Rq  )
knitr::kable(head(observations1))


#The next statement shows the first 10 triples in the cube, where the subject is a qb:Observation.

observations2Rq<-  paste( forsparqlprefix,
'
select *
where { ?s a qb:Observation ; ?p ?o .}
limit 1500
',
"\n"                               
)
observations2<- sparql.rdf(checkCube, observations2Rq)
knitr::kable(head(observations2))
#```

## Get cube components

#The cube components are shown in the next output.
#```{r, echo=TRUE}
componentsRq<- GetComponentSparqlQuery( forsparqlprefix, dsdName )
components<- as.data.frame(sparql.rdf(checkCube, componentsRq), stringsAsFactors=FALSE)
components$vn<- gsub("crnd-dimension:|crnd-attribute:|crnd-measure:","",components$p)
knitr::kable(components[,c("vn", "label")])


#The codelists are shown in the next output.
#```{r, echo=TRUE}
codelistsRq<- GetCodeListSparqlQuery( forsparqlprefix, dsdName )
codelists<- as.data.frame(sparql.rdf(checkCube, codelistsRq), stringsAsFactors=FALSE)
codelists$vn<- gsub("crnd-dimension:|crnd-attribute:|crnd-measure:","",codelists$dimension)
codelists$clc<- gsub("code:","",codelists$cl)
knitr::kable(codelists[,c("vn", "clc", "clprefLabel")])
```


The dimensions are shown in the next output.
```{r, echo=TRUE}
dimensionsRq <- GetDimensionsSparqlQuery( forsparqlprefix )
dimensions<- sparql.rdf(checkCube, dimensionsRq)
knitr::kable(dimensions)
```

Then the attributes as shown in the next output.
```{r, echo=TRUE}
attributesRq<- GetAttributesSparqlQuery( forsparqlprefix )
attributes<- sparql.rdf(checkCube, attributesRq)
knitr::kable(attributes)
```

## Get observations
And finally the SPARQL query for observations, showing only the first 10 observations.
```{r, echo=TRUE}
observationsRq<- GetObservationsSparqlQuery( forsparqlprefix, domainName, dimensions, attributes )
cat(observationsRq)
observations<- as.data.frame(sparql.rdf(checkCube, observationsRq ), stringsAsFactors=FALSE)
knitr::kable(observations[ 1:10 ,
   c(paste0(sub("crnd-dimension:|crnd-attribute:|crnd-measure:", "", dimensions), "value"),sub("crnd-dimension:|crnd-attribute:|crnd-measure:", "", attributes), "measure")])

```

## Get observations with labels

The SPARQL query for observations with labels for variables, showing only the first 10 observations.
```{r, echo=TRUE}
observationsDescriptionRq<- GetObservationsWithDescriptionSparqlQuery( forsparqlprefix, domainName, dimensions, attributes )
cat(observationsDescriptionRq)
observationsDesc<- as.data.frame(sparql.rdf(checkCube, observationsDescriptionRq ), stringsAsFactors=FALSE)
knitr::kable(observationsDesc[ 1:10 ,
     c(paste0(rep(sub("crnd-dimension:|crnd-attribute:|crnd-measure:", "", dimensions),each=3),
       c("label", "value", "IRI")),
       sub("crnd-dimension:|crnd-attribute:|crnd-measure:", "", attributes), "measure", "measureIRI"
       )]
       )

```



# Session information
```{r, echo=TRUE}
sessionInfo()
```

