---
title: "Create CDISC pilot tables"
author: " "
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    theme: united
  pdf_document:
    toc: true
    highlight: zenburn
  md_document:
    variant: markdown_github
---


# Setup

This script creates RDF data cubes from .csv files.

```{r, eval=TRUE}

library(rrdfancillary)
library(rrdfcdisc)
library(rrdfqb)
library(rrdfqbcrnd0)
```

All files are stored in the directory

```{r}
targetDir<- "../res-ttl"
(targetDir)
```

# Create from demo data PhUSE scripting program

The files are generated by a SAS program gen-res-sas/tab1x01.sas.  The input files are read
directly from .csv files instead of using the workbook, to avoid the
extra step with the workbook, as the input files may be updated by
running the SAS program.

```{r, eval=TRUE}
fig1x01ObsDataCsvFn<- file.path("../res-csv", "FIG1X01.csv")
fig1x01ObsData <- read.csv(fig1x01ObsDataCsvFn,stringsAsFactors=FALSE)


fig1x01MetaDataCsvFn<- file.path("../res-csv", "FIG1X01-Components.csv")
fig1x01MetaData <- read.csv(fig1x01MetaDataCsvFn,stringsAsFactors=FALSE)


fig1x01.cube.fn<- BuildCubeFromDataFrames(fig1x01MetaData, fig1x01ObsData )
cat("FIG1X01 cube stored as ", normalizePath(fig1x01.cube.fn), "\n")

targetFile<- file.path(targetDir,"CDISC-pilot-FIG1X01.ttl")

if (file.copy( fig1x01.cube.fn, targetFile, overwrite=TRUE)) {
   cat("RDF data cube copied to ", normalizePath(targetFile), "\n")
 }

```

Now look at the generated cubes by loading the turtle files. 

```{r, echo=TRUE, results='asis'}
dataCubeFile<- targetFile
```

The rest of the code only depends on the value of dataCubeFile.

```{r, echo=TRUE}
checkCube <- new.rdf(ontology=FALSE)  # Initialize
cat("Loading RDF data cube from ", normalizePath(dataCubeFile), "\n")
temp<- load.rdf(dataCubeFile, format="TURTLE", appendTo= checkCube)
summarize.rdf(checkCube)
```

## Get the values in the cube
First set values for accessing the cube.
```{r, echo=TRUE}
dsdName<- GetDsdNameFromCube( checkCube )
domainName<- GetDomainNameFromCube( checkCube )
forsparqlprefix<- GetForSparqlPrefix( domainName )
```

The next statement shows the first 10 triples in the cube.
```{r, echo=TRUE}
observations1Rq<- paste( forsparqlprefix,
'
select *
where {?s ?p ?o .}
limit 100
',
"\n"
)
observations1<- sparql.rdf(checkCube, observations1Rq  )
knitr::kable(head(observations1))
```

The next statement shows the first 10 triples in the cube, where the subject is a qb:Observation.

```{r, echo=TRUE}
observations2Rq<-  paste( forsparqlprefix,
'
select *
where { ?s a qb:Observation ; ?p ?o .}
limit 1500
',
"\n"                               
)
observations2<- sparql.rdf(checkCube, observations2Rq)
knitr::kable(head(observations2))
```

## Get cube components

The cube components are shown in the next output.
```{r, echo=TRUE}
componentsRq<- GetComponentSparqlQuery( forsparqlprefix, dsdName )
components<- as.data.frame(sparql.rdf(checkCube, componentsRq), stringsAsFactors=FALSE)
components$vn<- gsub("crnd-dimension:|crnd-attribute:|crnd-measure:","",components$p)
knitr::kable(components[,c("vn", "label")])
```

The codelists are shown in the next output.

```{r, echo=TRUE}
codelistsRq<- GetCodeListSparqlQuery( forsparqlprefix, dsdName )
codelists<- as.data.frame(sparql.rdf(checkCube, codelistsRq), stringsAsFactors=FALSE)
codelists$vn<- gsub("crnd-dimension:|crnd-attribute:|crnd-measure:","",codelists$dimension)
codelists$clc<- gsub("code:","",codelists$cl)
knitr::kable(codelists[,c("vn", "clc", "clprefLabel")])
```
The dimensions are shown in the next output.

```{r, echo=TRUE} 
dimensionsRq <- GetDimensionsSparqlQuery( forsparqlprefix )
dimensions<- sparql.rdf(checkCube, dimensionsRq)
knitr::kable(dimensions)
```

Then the attributes as shown in the next output.
```{r, echo=TRUE}
attributesRq<- GetAttributesSparqlQuery( forsparqlprefix )
attributes<- sparql.rdf(checkCube, attributesRq)
knitr::kable(attributes)
```

## Get observations
And finally the SPARQL query for observations, showing only the first 10 observations.
```{r, echo=TRUE}
observationsRq<- GetObservationsSparqlQuery( forsparqlprefix, domainName, dimensions, attributes )
cat(observationsRq)
observations<- as.data.frame(sparql.rdf(checkCube, observationsRq ), stringsAsFactors=FALSE)
knitr::kable(observations[ 1:10 ,
   c(paste0(sub("crnd-dimension:|crnd-attribute:|crnd-measure:", "", dimensions), "value"),sub("crnd-dimension:|crnd-attribute:|crnd-measure:", "", attributes), "measure")])

```

## Get observations with labels

The SPARQL query for observations with labels for variables, showing only the first 10 observations.
```{r, echo=TRUE}
observationsDescriptionRq<- GetObservationsWithDescriptionSparqlQuery( forsparqlprefix, domainName, dimensions, attributes )
cat(observationsDescriptionRq)
observationsDesc<- as.data.frame(sparql.rdf(checkCube, observationsDescriptionRq ), stringsAsFactors=FALSE)
knitr::kable(observationsDesc[ 1:10 ,
     c(paste0(rep(sub("crnd-dimension:|crnd-attribute:|crnd-measure:", "", dimensions),each=3),
       c("label", "value", "IRI")),
       sub("crnd-dimension:|crnd-attribute:|crnd-measure:", "", attributes), "measure", "measureIRI"
       )]
       )

```



# Session information
```{r, echo=TRUE}
sessionInfo()
```

