---
title: "poc-application-overview"
author: "mja@statgroup.dk"
date: "12 sep 2016"
output:
  word_document:
    toc: yes
  pdf_document:
    toc: yes
  html_document:
    highlight: pygments
    toc: yes
---

# Setup

The mermaid package included in DiagrammeR is used for drawing the flowcharts.
See (http://rich-iannone.github.io/DiagrammeR/graphviz_and_mermaid.html), (http://knsv.github.io/mermaid/#flowcharts-basic-syntax)

To make Word and PDF version the packages `htmltools`, `webshot` must be installed. `webshot` also needs phamtonjs, that can be installed using `webshot::install_phantomjs()`.

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
# install.packages(c("DiagrammeR","DiagrammeRsvg"))
# install.packages(c("webshot"))
library("DiagrammeR")
library("DiagrammeRsvg")
library("htmltools")
library("webshot")
```

Note: pdf generation does not work yet.

#Application process

Here is a high level overview of the proof of concept package

```{r pocdiag, echo=FALSE}
diagramSpec="
graph TB
subgraph ADaM data
A(ADaM)
end

subgraph Analysis results to RDF
A-->B[Analyse]
B-->C(Analysis Results as .csv)
C-->D[Convert to RDF datacube]
D-->E(Analysis Resuls as .ttl)
end

subgraph ADaM datasets results to RDF
A-->gA[Convert using D2RQ as .ttl]
gA-->aB(ADaM as .ttl)
end

subgraph Servering RDF
aB-->F
E-->F[SPARQL endpoint]
end

subgraph Presentation of results
F-->pA[Reporting]
pA-->pB(Tables as .html)
end

subgraph Servering HTML
pB-->pC[Webserver]
end

subgraph PoC Application
aF(.html and .js)-->pC
aG(configuration as .ttl)-->F
end

subgraph Browser
pC-->aA
F-->aA[Application]
end
"
## html_print(tagList(
##  tags$h1("Structure")
##  ,tags$pre(diagramSpec)
##  ,tags$div(class="mermaid",diagramSpec)
##  ,DiagrammeR()
##))
mermaid(diagramSpec)

```

# Example SPARQL query

```{r results='asis', engine='highlight', engine.path='C:/Program Files/Highlight/highlight.exe', engine.opts='-S n3 --inline-css', tidy=FALSE}
prefix rdfs:  <http://www.w3.org/2000/01/rdf-schema#>   
prefix qb: <http://purl.org/linked-data/cube#>   
prefix rrdfqbcrnd0: <http://www.example.org/rrdfqbcrnd0/>   
select (REPLACE( str(?ds), "^[^-]+-","") as ?shortname) ?dslabel ?underlyingData  
where {    
?ds  a qb:DataSet ;  
        rdfs:label                    ?dslabel ;  
        rrdfqbcrnd0:D2RQ-DataSetName  ?underlyingData .  
}
```    



# Testing SPARQL highlighting

Install highlight from (http://www.andre-simon.de/doku/highlight/en/highlight.php).
To avoid linebreak add two space at the end of each line (not good work around).
Another approach is to use http://pygments.org/languages/. Try searching for `pygments sparql`.

```{r results='asis', engine='highlight', engine.path='C:/Program Files/Highlight/highlight.exe', engine.opts='-S n3 --inline-css', tidy=FALSE}
select * where {   
  ?s ?p ?o   
  }   
```    
# Final stuff

Here is how to make the PDF file
```{r, eval=FALSE}
## rmarkdown::render('poc-application-overview.Rmd', c("pdf_document"), clean=TRUE)
rmarkdown::render('poc-application-overview.Rmd', c("word_document"), clean=TRUE)
```
