--
title: "Create TAB1X02_1 as RDF data cube"
author: "mja@statgroup.dk"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    theme: united
  pdf_document:
    toc: true
    highlight: zenburn
  md_document:
    variant: markdown_github
---


# Setup

This script creates RDF data cubes from .csv files.

```{r}
library(rrdfancillary)
library(rrdfcdisc)
library(rrdfqb)
library(rrdfqbcrnd0)
```

All files are stored in the directory
```{r}
targetDir<- "../res-ttl"
(targetDir)
```

# Load CSV files

```{r, eval=TRUE}
tab1x02ObsDataCsvFn<- file.path("../res-csv", "TAB1X02_1.csv")
tab1x02ObsData <- read.csv(tab1x02ObsDataCsvFn,stringsAsFactors=FALSE)

tab1x02MetaDataCsvFn<- file.path("../res-csv", "TAB1X02_1-Components.csv")
tab1x02MetaData <- read.csv(tab1x02MetaDataCsvFn,stringsAsFactors=FALSE)
```

# Create cube

```{r, eval=TRUE}
tab1x02.cube.fn<- BuildCubeFromDataFrames(tab1x02MetaData, tab1x02ObsData )
cat("TAB1X02 cube stored as ", normalizePath(tab1x02.cube.fn), "\n")
```
# Copy cube to destination directory

```{r, eval=TRUE}
targetFile<- file.path(targetDir,"CDISC-pilot-TAB1X02_1.ttl")

if (file.copy( tab1x02.cube.fn, targetFile, overwrite=TRUE)) {
   cat("RDF data cube copied to ", normalizePath(targetFile), "\n")
 }

```
# Get SPARQL query for observations

This is lifted from the example rrdfqbcrndex/vignettes/cube-from-workbook.Rmd.

```{r, eval=TRUE}
checkCube <- new.rdf()  # Initialize
temp<- load.rdf(targetFile, format="TURTLE", appendTo= checkCube)
summarize.rdf(checkCube)
```

## Get the values in the cube
First set values for accessing the cube.
```{r, echo=TRUE}
dsdName<- GetDsdNameFromCube( checkCube )
domainName<- GetDomainNameFromCube( checkCube )
forsparqlprefix<- GetForSparqlPrefix( domainName )
```

## Get cube components

The cube components are shown in the next output.
```{r, echo=TRUE}
componentsRq<- GetComponentSparqlQuery( forsparqlprefix, dsdName )
components<- as.data.frame(sparql.rdf(checkCube, componentsRq), stringsAsFactors=FALSE)
components$vn<- gsub("crnd-dimension:|crnd-attribute:|crnd-measure:","",components$p)
knitr::kable(components[,c("vn", "label")])
```

The codelists are shown in the next output.
```{r, echo=TRUE}
codelistsRq<- GetCodeListSparqlQuery( forsparqlprefix, dsdName )
codelists<- as.data.frame(sparql.rdf(checkCube, codelistsRq), stringsAsFactors=FALSE)
codelists$vn<- gsub("crnd-dimension:|crnd-attribute:|crnd-measure:","",codelists$p)
codelists$clc<- gsub("code:","",codelists$cl)
knitr::kable(codelists[,c("vn", "clc", "prefLabel")])
```


The dimensions are shown in the next output.
```{r, echo=TRUE}
dimensionsRq <- GetDimensionsSparqlQuery( forsparqlprefix )
dimensions<- sparql.rdf(checkCube, dimensionsRq)
knitr::kable(dimensions)
```

Then the attributes as shown in the next output.
```{r, echo=TRUE}
attributesRq<- GetAttributesSparqlQuery( forsparqlprefix )
attributes<- sparql.rdf(checkCube, attributesRq)
knitr::kable(attributes)
```

## Get observations
And finally the SPARQL query for observations, showing only the first 10 observations.
```{r, echo=TRUE}
observationsRq<- GetObservationsSparqlQuery( forsparqlprefix, domainName, dimensions, attributes )
cat(observationsRq)
observations<- as.data.frame(sparql.rdf(checkCube, observationsRq ), stringsAsFactors=FALSE)
knitr::kable(observations[ 1:10 ,
   c(paste0(sub("crnd-dimension:|crnd-attribute:|crnd-measure:", "", dimensions), "value"),sub("crnd-dimension:|crnd-attribute:|crnd-measure:", "", attributes), "measure")])

```

## Get observations with labels

The SPARQL query for observations with labels for variables, showing only the first 10 observations.
```{r, echo=TRUE}
observationsDescriptionRq<- GetObservationsWithDescriptionSparqlQuery( forsparqlprefix, domainName, dimensions, attributes )
cat(observationsDescriptionRq)
observationsDesc<- as.data.frame(sparql.rdf(checkCube, observationsDescriptionRq ), stringsAsFactors=FALSE)
knitr::kable(observationsDesc[ 1:10 ,
     c(paste0(rep(sub("crnd-dimension:|crnd-attribute:|crnd-measure:", "", dimensions),each=3),
       c("label", "value", "IRI")),
       sub("crnd-dimension:|crnd-attribute:|crnd-measure:", "", attributes), "measure", "measureIRI"
       )]
       )

```




# Session information
```{r, echo=TRUE}
sessionInfo()
```

