--
title: "Create TAB1X03 as RDF data cube"
author: "mja@statgroup.dk"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    theme: united
  pdf_document:
    toc: true
    highlight: zenburn
  md_document:
    variant: markdown_github
---

# Introduction
This script creates RDF data cubes from .csv files.

# Setup

All files are stored in the directory
```{r}
targetName<- "TAB1X03"
print(targetName)
targetDir<- "../res-ttl"
rqtargetDir<- "../sparql-rq"
print(targetDir)
targetFile<- file.path(targetDir, paste("CDISC-pilot-", targetName, ".ttl", sep=""))
print(targetFile)
targetObsrqFile<- file.path(targetDir, paste("CDISC-pilot-", targetName, "-observations", ".rq", sep=""))
print(targetObsrqFile)
target2DrqFile<- file.path(rqtargetDir, paste( tolower(targetName), ".rq", sep=""))
print(target2DrqFile)
```

# Create RDF data cube 

```{r, child='include_ttl.Rmd'}
```

# Generate SPARQL query for presentation

```{r}

rowdimensions<-c(
        "crnd-dimension:sitegr1",
        "crnd-dimension:siteid" 
    )

statdimensions<- c(
    "crnd-dimension:procedure",
    "crnd-dimension:factor"
    )

coldimensions<- c(
"crnd-dimension:trt01p",    
"crnd-dimension:ittfl",   
"crnd-dimension:efffl",   
"crnd-dimension:comp24fl"
    )


coldimval.rq<-
    paste( forsparqlprefix,
          "select distinct ", paste0( sub("crnd-dimension:", "?", coldimensions ), collapse=" ")," where {",
          "?s", " a ", "qb:Observation  ;",
          paste0( "    ", "qb:dataSet",  " ", "ds:", "dataset", "-", domainName, " ;", collapse="\n"),

          paste0( "    ", coldimensions, " ", sub("crnd-dimension:", "?", coldimensions ), ";", collapse="\n"),
          "}",
          "order by ", paste0( sub("crnd-dimension:", "?", coldimensions ), collapse=" ")
)
ll<- strsplit(coldimval.rq,split="\n")
cat(paste(seq(length(unlist(ll))),unlist(ll),collapse="\n"),"\n")

coldimval<- NULL
coldimval<- as.data.frame(sparql.rdf(checkCube, coldimval.rq ), stringsAsFactors=FALSE)
knitr::kable(coldimval)

colstack<- paste("zz",seq(nrow(coldimval)),sep="")

cc<-cbind(coldimensions,t(coldimval[1,]))
cat(paste( cc[,1],cc[,2], sep=" ", collapse="; \n")," ;\n")

```

```{r}
colz1.rq<- NULL



colz1.rq<-
    paste( forsparqlprefix,
          "select * where {",
          paste0( ## for each of column values
              paste0("?col", seq(nrow(coldimval)), "z1URI", " a qb:Observation  ;\n"),
              paste0( "    ", "qb:dataSet",  " ", "ds:", "dataset", "-", domainName, " ;\n", collapse="\n"),
              paste0( "    ", rowdimensions, " ", sub("crnd-dimension:", "?", rowdimensions ), ";\n", collapse=""),
              paste0( "    ", statdimensions, " ", sub("crnd-dimension:", "?", statdimensions), "z1", ";\n", collapse=""),
              paste0( coldimensions[1], " ", coldimval[,1], ";\n", sep=" "),
              paste0( coldimensions[2], " ", coldimval[,2], ";\n", sep=" "),
              paste0( coldimensions[3], " ", coldimval[,3], ";\n", sep=" "),
              paste0( coldimensions[4], " ", coldimval[,4], ";\n", sep=" "),
              paste0( " crnd-measure:measure ",  "?col", seq(nrow(coldimval)), "z1", sep=" "),
              ".", 
              sep="\n", collapse="\n"
              ),
"optional {",                     
paste0("",
sub("crnd-dimension:", "?", rowdimensions), " ", "skos:prefLabel", " ", sub("crnd-dimension:", "?", rowdimensions), "value", ".", collapse="\n"),                     
paste0(  rowdimensions , " ",
           "rdfs:label", " ", sub("crnd-dimension:", "?", rowdimensions), "label" ,
           " . ", 
           collapse="\n"),
## paste0(" ",
## sub("crnd-dimension:", "?", statdimensionsz1z2), " ", "skos:prefLabel", " ", sub("crnd-dimension:", "?", statdimensionsz1z2), "value", ".", collapse="\n"),                     
## paste0( sub("crnd-dimension:", "?", statdimensionsz1z2) , " ",
##            "rdfs:label", " ", sub("crnd-dimension:", "?", statdimensionsz1z2), "label" ,
##            " . ", 
##        collapse="\n"),
"}",
"\n{ select  ?procedurez1 where {
   code:procedure skos:hasTopConcept ?procedurez1 .
   filter ( ?procedurez1 != code:procedure-percent )
   } }
",
"}"
)
cat(colz1.rq,file=target2DrqFile)
ll<- strsplit(colz1.rq,split="\n")
cat(paste(seq(length(unlist(ll))),unlist(ll),collapse="\n"),"\n")

colz1obs<- NULL
colz1obs<- as.data.frame(sparql.rdf(checkCube, colz1.rq ), stringsAsFactors=FALSE)
knitr::kable(colz1obs)

```


