---
title: "Create TAB5X01 as RDF data cube"
author: "mja@statgroup.dk"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    theme: united
  pdf_document:
    toc: true
    highlight: zenburn
  md_document:
    variant: markdown_github
---

# Introduction
This script creates RDF data cubes from .csv files.

# Setup

All files are stored in the directory
```{r}
targetName<- "TAB5X01"
targetDir<- "../res-ttl"
rqtargetDir<- "../sparql-rq"
targetFile<- file.path(targetDir, paste("CDISC-pilot-", targetName, ".ttl", sep=""))
targetObsrqFile<- file.path(targetDir, paste("CDISC-pilot-", targetName, "-observations", ".rq", sep=""))
target2DrqFile<- file.path(rqtargetDir, paste( tolower(targetName), ".rq", sep=""))
```

The rest of the generation is made by the script `include_ttl.Rmd`.

```{r, child='include_ttl.Rmd'}
```

The next piece can be made more generic and expressed shorter.

The structure of the rectangular dataset is given from the column dimensions.
The row dimension is the the remaining dimensions excluding the two dimensions representings statistics.
So the row dimensions can be derived from the cube structure. This also makes sense as the
column dimension is the same for more tables than the row dimensions.

The levels for the columns can also be determined from the datacube, including the renaming of columns.

```{r}

rowdimensions<- NULL
rowdimensions<-c(
"crnd-dimension:saffl",
"crnd-dimension:trtemfl",
"crnd-dimension:aesoc",
"crnd-dimension:aedecod"
    )

statdimensions<- NULL
statdimensions<- c(
    "crnd-dimension:procedure",
    "crnd-dimension:factor"
    )

coldimension<- NULL
coldimension<- c("crnd-dimension:trta" )

colvalues<- NULL
colvalues<- c( "code:trta-Placebo",
"code:trta-Xanomeline_Low_Dose",
"code:trta-Xanomeline_High_Dose"
)               

statdimensionsz1z2<- NULL
statdimensionsz1z2<- c(paste0(statdimensions,"z1"), paste0( statdimensions, "z2"))
statdimensionsz1z2<- c(paste0(statdimensions,"z1"))
colz1.rq<- NULL
colz1.rq<-     paste( forsparqlprefix,
"select * where {",
                     paste0( ## for each of column values
"?col", seq(length(colvalues)), "z1URI", " a qb:Observation  ;", "\n",
paste0( "    ", "qb:dataSet",  " ", "ds:", "dataset", "-", domainName, " ;", "\n", collapse="\n"),
paste0( "    ", rowdimensions, " ", sub("crnd-dimension:", "?", rowdimensions ), ";", collapse="\n"),
paste0( "    ", statdimensions, " ", sub("crnd-dimension:", "?", statdimensions), "z1", ";", collapse="\n"),
" ",
"    ", coldimension, " ", colvalues, " ;",    
" ", "\n",                     
"    crnd-measure:measure ",  "?col", seq(length(colvalues)), "z1", " .", "\n",
"optional{", "\n",
"?col", seq(length(colvalues)), "z2URI", " a qb:Observation  ;",
paste0( "    ", "qb:dataSet",  " ", "ds:", "dataset", "-", domainName, " ;", collapse="\n"),
paste0( "    ", rowdimensions, " ", sub("crnd-dimension:", "?", rowdimensions ), ";", collapse="\n"),
"      crnd-dimension:procedure code:procedure-percent; \n",
"    crnd-dimension:factor code:factor-proportion;  \n",
"    crnd-attribute:denominator ",  "?col", seq(length(colvalues)), "z2denominator; \n", 
"    ", coldimension, " ", colvalues, " ;",    
# "    crnd-dimension:procedure ?procedurez2 ; \n",
# "    crnd-dimension:factor ?factorz2 ;  \n",
" ",                     
"    crnd-measure:measure ",  "?col", seq(length(colvalues)), "z2", " . \n",
    "?col", seq(length(colvalues)), "z1URI", " ", "crnd-dimension:procedure", " ", "code:procedure-count", " .",    "\n",
"}", " .",
                         collapse="\n"),
"optional {", "\n",                     
paste0("",
sub("crnd-dimension:", "?", rowdimensions), " ", "skos:prefLabel", " ", sub("crnd-dimension:", "?", rowdimensions), "value", ".", collapse="\n"),                     
paste0(  rowdimensions , " ",
           "rdfs:label", " ", sub("crnd-dimension:", "?", rowdimensions), "label" ,
           " . ", 
           collapse="\n"),
## paste0(" ",
## sub("crnd-dimension:", "?", statdimensionsz1z2), " ", "skos:prefLabel", " ", sub("crnd-dimension:", "?", statdimensionsz1z2), "value", ".", collapse="\n"),                     
## paste0( sub("crnd-dimension:", "?", statdimensionsz1z2) , " ",
##            "rdfs:label", " ", sub("crnd-dimension:", "?", statdimensionsz1z2), "label" ,
##            " . ", 
##        collapse="\n"),
 "\n",   "}",
"
{ select  ?procedurez1 where {
   code:procedure skos:hasTopConcept ?procedurez1 .
   filter ( ?procedurez1 != code:procedure-percent )
   } }
",
"}"
)
cat(colz1.rq,file=target2DrqFile)
ll<- strsplit(colz1.rq,split="\n")
cat(paste(seq(length(unlist(ll))),unlist(ll),collapse="\n"),"\n")

colz1obs<- NULL
colz1obs<- as.data.frame(sparql.rdf(checkCube, colz1.rq ), stringsAsFactors=FALSE)
knitr::kable(colz1obs)

```


